---
title: "Social Network Analysis (SNA) Part2"
author: "Nik Bear Brown"
output:
  html_document: default
  word_document: default
---

In this lesson we'll learn the how to implement and analyze Social network analysis (SNA) in R.


# Additional packages needed
 
To run the code you may need additional packages.

* If necessary install the followings packages.

`install.packages("igraph");`  
`install.packages("tools");`  
`install.packages("visNetwork");`  


```{r}
library(igraph)
library(tools)
library(visNetwork)
```


# Data

We will be creating graphs with the library(igraph).

```{r}
# for reproducibility of graphs plots (plot.igraph uses random numbers)
set.seed(3333)
# create an example graph
D <- read.table(header=T,text=
                  'from to
A B
A C
C D
C F
C E
D E
D F
E F')
g1 <- graph.data.frame(D,directed=F)
# plot the original graph
plot(g1)
# find all the largest cliques (returns a list of vector of vertiex ids)
a <- largest.cliques(g1)
# let's just take the first of the largest cliques
# (in this case there's just one clique)
clique1 <- a[[1]]
# subset the original graph by passing the clique vertices
g2 <- graph.full(length(clique1))
V(g2)$name <- V(g1)$name[clique1]
# plot the clique
plot(g2)
```

In addition we will be using social relations among Renaissance Florentine families (person aggregates) collected by John Padgett from historical documents. The two relations are business ties (PADGB - specifically, recorded financial ties such as loans, credits and joint partnerships) and marriage alliances (PADGM). See http://moreno.ss.uci.edu/data.html#padgett

```{r}
# Studying marriage ties among Renaissance Florentine families
data_url <- 'http://nikbearbrown.com/YouTube/MachineLearning/M12/padgett-cleaned.txt'
padgett.cleaned <- read.table(url(data_url), quote="\"")
```

# Cliques and Motifs among Renaissance Florentine families

Find and plot all the largest cliques seperately. Find any significant motifs.   

```{r}
florentine_marriage_ties <- padgett.cleaned[1:16,]
rownames(florentine_marriage_ties) <- c("ACCIAIUOLI", "ALBIZZI", "BARBADORI", 
                                        "BISCHERI", "CASTELLANI", "GINORI", 
                                        "GUADAGNI", "LAMBERTES","MEDICI", 
                                        "PAZZI", "PERUZZI", "PUCCI", 
                                        "RIDOLFI", "SALVIATI", "STROZZI", 
                                        "TORNABUON")
colnames(florentine_marriage_ties) <- rownames(florentine_marriage_ties)
florentine_graph <- 
  graph_from_adjacency_matrix(as.matrix(florentine_marriage_ties))                
plot(florentine_graph, layout=layout.fruchterman.reingold,
     main = "Marriage ties among Florentine families",
     vertex.label=V(florentine_graph)$name, vertex.size=25, 
     vertex.color="blue", vertex.frame.color="white",
     vertex.label.color="black", vertex.label.cex=1,
     edge.arrow.size=0.25, edge.width=0.25)
count_motifs(florentine_graph, size = 3)                
count_motifs(florentine_graph, size = 4)
# find and plot all the largest cliques seperately
florentine_graph_undirected <- as.undirected(florentine_graph)
# following code taken from
# http://stackoverflow.com/questions/26222659/identifying-cliques-in-r 
all <- largest_cliques(florentine_graph_undirected)
all
clique_graph1 <- make_full_graph(length(all[[1]]))
V(clique_graph1)$name <- V(florentine_graph_undirected)$name[all[[1]]]
plot(clique_graph1)
clique_graph2 <- make_full_graph(length(all[[2]]))
V(clique_graph2)$name <- V(florentine_graph_undirected)$name[all[[2]]]
plot(clique_graph2)
clique_graph3 <- make_full_graph(length(all[[3]]))
V(clique_graph3)$name <- V(florentine_graph_undirected)$name[all[[3]]]
plot(clique_graph3)
# plot all largest cliques
# following code taken from R and Data Mining: Examples and Case Studies 
# by Yanchang Zhao, page 118
florentine_largest_cliques <- largest_cliques(florentine_graph_undirected)
length(florentine_largest_cliques)
colorbar <- rainbow(length(florentine_largest_cliques) + 1)
for (i in 1:length(florentine_largest_cliques)) {
  V(florentine_graph_undirected)[florentine_largest_cliques[[i]]]$color <- colorbar[i+1]
}
plot(florentine_graph_undirected, mark.groups=florentine_largest_cliques,
     vertex.size=0.3, vertex.label.cex=1, edge.color=rgb(1,0.5,1,1),
     main = "Largest Cliques")
# plot all maximal cliques
# following code taken from R and Data Mining: Examples and Case Studies 
# by Yanchang Zhao, page 117
florentine_max_cliques <- max_cliques(florentine_graph_undirected)
length(florentine_max_cliques)
colorbar <- rainbow(length(florentine_max_cliques) + 1)
for (i in 1:length(florentine_max_cliques)) {
  V(florentine_graph_undirected)[florentine_max_cliques[[i]]]$color <- colorbar[i+1]
}
plot(florentine_graph_undirected, mark.groups=florentine_max_cliques,
     vertex.size=0.3, vertex.label.cex=1, edge.color=rgb(1,0.5,1,1),
     main = "Maximal Cliques")
# Calculate PageRank of the vertices
ranks_of_families <- page_rank(florentine_graph)$vector
ranks_of_families
# most important or influential family
which.max(ranks_of_families) # 9 is the index of the vertex

```


# Resources   

* [Network visualization in R with the igraph package](https://rulesofreason.wordpress.com/2012/11/05/network-visualization-in-r-with-the-igraph-package/)    

* [Making prettier network graphs with sna and igraph via @rbloggers](http://www.r-bloggers.com/making-prettier-network-graphs-with-sna-and-igraph/)       

* [igraph R manual pages](http://igraph.org/r/doc/)    

# References

The data, R code and lessons are based upon:   

Graph theory/data structures:   
* http://math.tut.fi/~ruohonen/GT_English.pdf   
* http://www.cl.cam.ac.uk/teaching/1011/PrincComm/slides-lpr/graph_theory_1-11.pdf   
* http://www.researchgate.net/publication/228300013_Graph_Theory_A_Primer_for_Using_R_Visualization_Techniques_in_the_Applications_of_the_Adjacency_Matrix   
* http://www.boost.org/doc/libs/1_59_0/libs/graph/doc/graph_theory_review.html   
SNA:   
* http://files.meetup.com/1406240/sna_in_R.pdf   
* http://www2.unb.ca/~ddu/6634/Lecture_notes/Lec1_intro_handout.pdf    
* http://www.faculty.ucr.edu/~hanneman/nettext/   
* http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.372.1960&rep=rep1&type=pdf   
* http://www.rdatamining.com/examples/social-network-analysis   

igraph:   
* http://statmath.wu.ac.at/research/friday/resources_WS0708_SS08/igraph.pdf   
* http://blog.revolutionanalytics.com/2014/11/a-look-at-the-igraph-package.html   
* http://www.r-bloggers.com/igraph-and-sna-an-amateurs-dabbling/    
* http://www.r-bloggers.com/going-viral-with-rs-igraph-package/    
* https://cran.r-project.org/web/packages/igraph/igraph.pdf    

Other famous packages for SNA:    
* http://www.r-bloggers.com/must-have-r-packages-for-social-scientists/   
* https://cran.r-project.org/web/views/SocialSciences.html   
* https://cran.r-project.org/web/packages/sna/sna.pdf   
* https://cran.r-project.org/web/packages/RSiena/RSiena.pdf   
* https://cran.r-project.org/web/packages/network/network.pdf   
* https://www.bioconductor.org/packages/release/bioc/manuals/graph/man/graph.pdf   
* http://www.statnet.org/    

In-depth SNA tutorials:   
* http://sna.stanford.edu/rlabs.php   
* http://www.stats.ox.ac.uk/~snijders/sna_course.htm   
* http://www.shizukalab.com/toolkits  

Sample projects:   

* http://www.orgnet.com/cases.html 

Motifs:   
* http://igraph.org/r/doc/motifs.html   
* https://en.wikipedia.org/wiki/Network_motif   
* http://www.cs.columbia.edu/4761/notes07/chapter8.2-topology.pdf   
* https://sites.google.com/site/networkanalysisacourse/schedule/networkmotifs   
Cliques:   
* http://igraph.org/r/doc/cliques.html   
* http://faculty.ucr.edu/~hanneman/nettext/C11_Cliques.html   
* https://courses.cs.washington.edu/courses/cse527/01au/oct25/oct25.html    
* ttp://www.mathcove.net/petersen/lessons/get-lesson?les=29    
* http://news.stanford.edu/news/2014/november/cliques-high-school-110514.html    
PageRank:    
* http://igraph.org/r/doc/page_rank.html    
* http://ilpubs.stanford.edu:8090/422/1/1999-66.pdf    
* http://www.cs.princeton.edu/~chazelle/courses/BIB/pagerank.htm    
* http://www.stat.cmu.edu/~ryantibs/datamining/lectures/03-pr-marked.pdf   
* http://smallstats.blogspot.com/2014/04/from-random-walks-to-personalized.html    
* http://blog.revolutionanalytics.com/2014/12/a-reproducible-r-example-finding-the-most-popular-packages-using-the-pagerank-algorithm.html    
* Mining Massive Datasets on Coursera - Week 1, Videos 5 through 11 explain PageRank elegantly. The course maybe unavailable (or archived) by the time this module is out   

Dataset:   
* http://moreno.ss.uci.edu/data.html#padgett   
* http://home.uchicago.edu/~jpadgett/papers/unpublished/maelite.pdf   

Other SNA:  
* http://www.r-bloggers.com/experiments-with-igraph/   
*  http://cran.us.r-project.org/doc/contrib/Zhao_R_and_data_mining.pdf   
 
```



```
