---
title: "Professor Bear - R Data structures"
output:
  html_document: default
  word_document: default
---


**Table of Contents**  
[Section 1.1: Data structures in R](#data-structures-in-r)  
[Section 1.1.1: Vectors](#vectors)  
[Section 1.1.1: Lists](#lists)  
[Section 1.1.1: Arrays](#arrays)  
[Section 1.1.1: Matrices](#matrices)  
[Section 1.1.1: Data frames](#data-frames)  

##1.1: Data structures in R  

Data Structures are the programmatic way of storing data so that data can be used efficiently. One can think of a data structure as a container for data. 

##1.1.1: Vectors  

Usually when you create variables in R, you create **vectors**. A vector is simply a set of elements *of the same class* (e.g. character, numeric, integer, or logical -as in True/False). It is the basic data structure in R. Most commonly, you will use the `c()` function (c stands for concatenate) to create vectors.
:

```{r}
v1 <- c(1,2,3,4,5) #creates a numeric vector
v1
v2 <- c(1L, 2L) #creates an integer vector
v2
v3 <- c(TRUE, FALSE) #creates a logical vector
v3
v4 <- c("a", "b", "c") #creates a character vector
v4
v5 <- c(1+0i, 3+5i) #creates a complex vector 
v5

```

*Operations on vectors*

Once you have a vector (or a list of numbers) in memory most basic operations are available. This makes R very powerful.

```{r}
v1 
v1 + 5 # add 5 to each of the numbers
v1*3 # Multiply each of the numbers by 3
v1/3 # Divide each of the numbers by 3
log(v1) # Take the log of each of the numbers
```

We can even add vectors to vectors.

```{r}
a <- c(1,2,3)
a
b <-  c(1,2,3,4,5)
b
a+b
```

If the lengths of the vectors differ then you may get an error message, a warning message and unpredictable results. It is best if they are the same length.

```{r}
a <- c(1,2,3)
a
length(a)
b <-  c(3,4,5)
b
length(b)
a+b
```


If you mix in a vector elements that are of a different class (for example numerical and logical), R will **coerce** to the minimum common denominator, so that every element in the vector is of the same class. 

```{r}
b <- c(33, "bear")
class(b)
```


Note what happens when a vector is placed in another vector.

```{r}
x <- c(2, 3, 5)
x
class(x)
y <- c(6, x, 2)
y
class(y)
```



##1.1.1: Lists  

A list is a generic data structure containing other objects. Unlike a vector it allows us to create a mixed data sequence. 

```{r}
w <-  c(1, 2, 3) 
w
class(w)
x <- c("a", "b", "c", "d", "e") 
x
class(x)
y <- c(TRUE, FALSE, TRUE, FALSE) 
y
z <- list(w, y, x, 33, "bear", FALSE) # create a mixed data type list
z
class(z)

```

*List Slicing*

We use the single square bracket "[]" operator to access elements at the first level of a list. The index 1 is the first element. We use the doubkle square bracket "[]" operator to access elements at the second level, etc.

```{r}
z
z[1] # First element 
z[[2]] # cond element 
z[[2]][1] # First element of the second element 

```



##1.1.1: Arrays  

Vectors are one-dimensional arrays in R and matrices are two-dimensional arrays in R. We can create n-dimensional arrays as a set of stacked matrices of identical dimensions. This will be discuss in the [matrices](#matrices) section below. The term arrays is discussed here becuase most programming languages use the term array. The reason R uses the term vector, is the basic operations can be applied to vectors whereas most programming languages use functions and loops to apply operations to arrays.  

##1.1.1: Matrices  

Matrices are n-dimensional vectors (usually two-dimensional). We build matrices from vectors (one-dimensional arrays). You can create a matrix in two ways. By using the command *matrix*.
```{r}
x<-matrix(c(1,2,3,4,5,6), nrow = 2, ncol = 3)
x
class(x)
```

Or using *cbind()* or *rbind()* to add columns or rows to a vectors.

```{r}
x <- c(1, 2, 3)         # Creates a vector `x' of 3 values.
x
class(x)
y <- c(55, 33, 11)         # Creates another vector `y' of 3 values.
y
class(y)
a<-rbind(x, y) # Creates a 2 x 3 matrix.  Note that the rows are appended
a
class(a)
b<-cbind(x, y)   # Creates a 3 x 2 matrix.  Note that the columns are appended
b
class(b)
```

*Matrix element access*

As with vectors, square brackets extract specific values from a matrix but more values are used within the brackets seperated by commas for each diminasion.

 
```{r}
b
b[,2] # Extracts the second column
b[1,]  # Extracts the first row
b[1,2] # Extracts element in the first row and the second column
b[1,2]<=55 # Recplaces the element in the first row and the second column with 55
b
b[1,] <- c(3,3) # Replaces the first row 
b
```


*Matrix operations and functions*

R supports a variety of matrix functions, including: det(), which returns the matrix's determinant; t(), which transposes the matrix; solve(), which inverts the the matrix; dim() command returns the dimensions of your matrix. 
```{r}
b
dim(b)
```

##1.1.1: Data frames  

A data frame is used for storing data tables of mixed data type. Typically, data from an excel sheet will be imported in to R as a data frame. We will use the built-in data set “InsectSprays” to discuss data frames.

```{r}
data(InsectSprays)
names(InsectSprays)
head(InsectSprays,3)
dim(InsectSprays)
nrow(InsectSprays) 
ncol(InsectSprays) 
levels(InsectSprays$spray)
summary(InsectSprays)
```

##1.17: Further resources  

[LearnR](https://youtu.be/p3i7Kz6C_-4?list=PLFAYD0dt5xCwDNFdrqeNoU9t-nhAWkbKe)  


[Try R @codeschool](http://tryr.codeschool.com)  

[Datacamp R Tutorials](https://www.datacamp.com/)   


[rstudio online learning](https://www.rstudio.com/online-learning/)  
