---
title: "Introduction to Functional Programming in R"
author: "Daniel Emaasit"
date: "February 27, 2016"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
suppressPackageStartupMessages(library(pryr)) # install.packages("pryr")
```


Note: This is from *Functional Programming in R*  [https://github.com/Emaasit/Functional-Programming-in-R](https://github.com/Emaasit/Functional-Programming-in-R) by Daniel Emaasit   

## Introduction
R is a functional programming language, meaning that it has tools for creating and manipulating functions.

## First class Functions
R has first class functions. You can do anything with functions that you can do with vectors, like:  
* Assign them to variables
* Store them in lists
* Pass them as arguments
* Create them inside functions
* Return them as a result of a function

## Functional Components
All R functions are made up of 3 components:

* the body
* the arguments/formals
* the environment

```{r show_func_components}
## By printing the function
f <- function(n, mean, sd){
  ## Generate n random normal observations with mean = mean and std dev = sd
  rnorm(n, mean, sd)
}

print(f) ## if environment is not shown, it means it was created in the Global Env
body(f)
formals(f)
environment(f)
attributes(f)
class(f)
f(10, 1, 1)
```

### Primitive Functions
Functions (only found in base package) whose `formals()`, `body()`, and `environment()` are all `NULL`. Like:
```{r primitive_funcs}
sum
formals(sum)
body(sum)
environment(sum)
```

## Scoping
The set of rules that governs how R looks up the values of a symbol.
Two types of scopint:

* Lexical Scoping: Looks up symbol values based on how functions were nested when they were created, not how they are nested when they are called
* Dynamic Scoping: Used in select functions to save typing during interactive analysis

Four basic principles of R's Lexical Scoping

* name masking
* functions vs variables
* a fresh start
```{r fresh_start}
j <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  print(a)
}

j()  
```
* dynamic lookup

**Closures**: Functions created by other functions

## Every Operation is a function call



## Function Arguments


## Special Calls


## Return Values