---
title: "Functionals in R"
author: "Daniel Emaasit"
date: "February 28, 2016"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

Note: This is from *Functional Programming in R*  [https://github.com/Emaasit/Functional-Programming-in-R](https://github.com/Emaasit/Functional-Programming-in-R) by Daniel Emaasit   

## Functionals
Functions that take a function as input and returns a vector as output. Examples:

* lapply
* sapply + vapply (vector output)
* mapply + Map (multiple inputs)
* apply (matrices & arrays)
* tapply

These functionals are already implemented in base R.

#### Uses of Functionals

* Commonly used as an alternative to **For Loops**.
* For encapsulating common data manipulation tasks like split-apply-combine

#### Advantages using Functionals

* Reduce bugs
* Functionals implemented in base R are efficient & fast

### lapply (for lists)
`lapply()` takes a function, applies it to each element in a list, & returns a list.
```{r my_first_functional}
input_list <- as.list(mtcars)
output_list <- lapply(input_list, length)
unlist(output_list)

mtcars[] <- lapply(mtcars, function(x) x / mean(x))
```

#### Looping Pattern 1: Over elements
Looping over the elements in a list
```{r loop_elements, eval = FALSE}
lapply(xs, function(x) {})
```

#### Looping Pattern 2: Over numeric indices
Looping over the numeric indices in a list
```{r loop_numeric_indices, eval = FALSE}
lapply(seq_along(xs), function(x) {})
```

#### Looping Pattern 3: Over the names
Looping over the names in a list
```{r loop_names, eval = FALSE}
lapply(names(xs), function(nm) {})
```

### sapply & vapply (vector outputs)
Functionals that take a function, apply it to every element in a list, and return an atomic vector.

* `sapply()` guesses while `vapply()` takes an additional argument for the output type
* `vapply()` is better suited for use inside functions.

```{r vector_output}
df <- data.frame(x = 1:10, y = Sys.time() + 1:10)
sapply(df, class)
# vapply(df, class, character(2))
```

### mapply & Map (multiple inputs)
Used when you have two or more lists (or data frames) that you need to process in parallel.
```{r Map_example}
# Generate some sample data
xs <- replicate(n = 5, expr = runif(10), simplify = FALSE)
ws <- replicate(n = 5, expr = rpois(10, 5) + 1, simplify = FALSE)

Map(function(xs, ws) {weighted.mean(xs, ws)}, xs, ws)
```

#### Parallelisation (mclapply + mcMap)
Since we can compute each element in any order, it's easy to dispatch tasks to different cores, and compute them in parallel using the **parallel** package.
```{r parralelisation}
library(parallel)
system.time(mclapply(1:1000, sqrt, mc.cores = 4))
system.time(lapply(1:1000, sqrt))
```

### apply (for matrices & arrays)
`apply()` is a variant of lapply for working hig-order dimensional data objects
```{r apply_example}
m <- matrix(1:100, nrow = 10)
apply(m, 1, mean) ## MARGINS, 1 = rows & 2 = columns
apply(m, 2, mean)
```

### tapply (for ragged arrays)
Ragged arrays are arrays where each row can have a different number of columns. `apply()` is useful for sumarizing a data set.
```{r tapply_example}
## Generate some ragged data
pulse <- round(rnorm(n = 22, mean = 70, sd = 10 / 3) + rep(c(0, 5), c(10, 12)))
group <- rep(c("A", "B"), c(10, 12))
split(pulse, group)

tapply(pulse, group, length)
tapply(pulse, group , mean)
```

### the plyr package

### Reduce
Reduce a vector to a single value
```{r reduce_example}
Reduce(sum, 1:100)

## Find the values that occur in each element in this list
l <- replicate(n = 5, sample(x = 1:10, size = 15, replace = TRUE), simplify = FALSE)
Reduce(intersect, l)
```

### Predicate Functionals (Filter, Find, Position)
* Predicates are functions that return a single TRUE or FALSE (e.g. is.character)
* Predicate Functionals applies a predicate to each element of a list or data frame.
```{r predicates}
df <- data.frame(x = 1:3, y = c("a", "b", "c"))

Filter(is.factor, df)
Find(is.factor, df)
Position(is.factor, df)
```

### Mathematical Functionals

