---
title: "Professor Bear - Importing Data in R"
author: "Bear"
date: "September 1, 2018"
output:
  word_document: default
  html_document: default
---

The first step in data analysis is getting the data in to R. Small datasets often come in the form of Excel  (`.xls`), a comma delimited (Comma-Separated
Value/CSV or `.csv`) or tab delimited 
(Tab-Separated Value/TSV/TXT e.g. `.txt`) files. 

## Paths and the Working Directory

First one needs to identify your *working directory*. This is the directory or folder in which R will save or look for files by default. 
As a reminder, you can see your working directory by typing:

```{r}
getwd()
```

You can also change your working directory using the function `setwd()`. Or you can change it through RStudio by clicking on "Session". 

## Functions to read in data into R

The are several functions in base R that are available for reading data.

## read.csv

read.csv reads a file in csv format and creates a data frame from it, with cases corresponding to lines and variables to fields in the file.

```{r,eval=FALSE}
?read.csv
```

Type `?read.csv` to learn how to use its arguments.


```{r,eval=FALSE}
read.csv(file, header = TRUE, sep = ",", quote = "\"",
         dec = ".", fill = TRUE, comment.char = "", ...)
```

Using read.csv to load some data.

```{r read.csv}
# Load our data using read.csv

data_url <- 'https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/master/csv/psych/galton.csv'
galton <- read.csv(url(data_url))
class(galton)
head(galton)
summary(galton)
```

## read.table

read.table reads a file in table format and creates a data frame from it, with cases corresponding to lines and variables to fields in the file.

```{r,eval=FALSE}
?read.table
```

Type `?read.table` to learn how to use its arguments.

```{r,eval=FALSE}
read.table(file, header = FALSE, sep = "", quote = "\"'",
           dec = ".", numerals = c("allow.loss", "warn.loss", "no.loss"),
           row.names, col.names, as.is = !stringsAsFactors,
           na.strings = "NA", colClasses = NA, nrows = -1,
           skip = 0, check.names = TRUE, fill = !blank.lines.skip,
           strip.white = FALSE, blank.lines.skip = TRUE,
           comment.char = "#",
           allowEscapes = FALSE, flush = FALSE,
           stringsAsFactors = default.stringsAsFactors(),
           fileEncoding = "", encoding = "unknown", text, skipNul = FALSE)
```

Using read.table to load some data.

```{r read.table}
# Load our data using read.table
# Balloons Data Set
data_url <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/balloons/adult+stretch.data'
balloons <- read.table(url(data_url))
class(balloons)
head(balloons)
summary(balloons)
```

Whoops, what happened?  Look at the [Balloons Data Set](https://archive.ics.uci.edu/ml/machine-learning-databases/balloons/adult+stretch.data)   

```{r read.table.ggod}
balloons <- read.table(url(data_url), sep = ",")
class(balloons)
head(balloons)
summary(balloons)
```


## read.delim

read.delim reads a file in tab delimited table format and creates a data frame from it, with cases corresponding to lines and variables to fields in the file.

```{r, eval = FALSE}
# set your working directory - normally where you data are
setwd('path/to/your/data')
data = read.delim('data.file',
                  header = TRUE, 
                  sep = '\t')
```

Type `?read.delim` to learn what the `header` and `sep` arguments do.

```{r,eval=FALSE}
?read.delim
```


```{r,eval=FALSE}
read.delim(file, header = TRUE, sep = "\t", quote = "\"",
           dec = ".", fill = TRUE, comment.char = "", ...)
```

## Quiz - load some data with read.delim

Find some data on the [UC Irvine Machine Learning Repository](http://archive.ics.uci.edu/ml/) and load it with read.delim


```












```



