---
title: "Text Processing in R"
author: "Professor Bear"
output:
  html_document: default
  word_document: default
---

In this lesson we'll learn text processing and regular expressions in R.


# Additional packages needed
 
* If necessary install the followings packages.

`install.packages("stringr");`   


```{r}
library(stringr)
```

# Text Processing in R

Lets start with an example string.

```{r}
my_string <- "Find a goal you are passionate about. Then pursue it. Relentlessly."
my_string
lower_string <- tolower(my_string)
lower_string
my_string_vector <- str_split(my_string, " ")
my_string_vector

```

Here's a quick cheat-sheet on [string manipulation functions in R](http://www.statmethods.net/management/functions.html), from Quick-R’s list of String Functions.

```{r, eval=FALSE}
substr(x, start=n1, stop=n2)
grep(pattern,x, value=FALSE, ignore.case=FALSE, fixed=FALSE)
gsub(pattern, replacement, x, ignore.case=FALSE, fixed=FALSE)
gregexpr(pattern, text, ignore.case=FALSE, perl=FALSE,
fixed=FALSE)
strsplit(x, split)
paste(…, sep="", collapse=NULL)
sprintf(fmt, …)
toupper/tolower(x)
nchar(x)
```

# Regular Expressions and Grep

In theoretical computer science and formal language theory, a [regular expression](https://en.wikipedia.org/wiki/Regular_expression) (abbreviated regex or regexp and sometimes called a rational expression) is a sequence of characters that define a search pattern, mainly for use in pattern matching with strings, or string matching, i.e. "find and replace"-like operations. The concept arose in the 1950s, when the American mathematician Stephen Kleene formalized the description of a regular language, and came into common use with the Unix text processing utilities ed, an editor, and [grep](https://en.wikipedia.org/wiki/Grep) (global regular expression print), a filter.

grep is a command-line utility for searching plain-text data sets for lines matching a regular expression. Grep was originally developed for the Unix operating system, but is available today for all Unix-like systems and is built in to languages like python and Perl.

# Regular Expressions Examples

Basic regex syntax

```
.	Normally matches any character except a newline.  

When you match a pattern within parentheses, you can use any of $1, $2, ... later to refer to the previously matched pattern.	

+	Matches the preceding pattern element one or more times.  	
?	Matches the preceding pattern element zero or one times.  	
*	Matches the preceding pattern element zero or more times.  
|	Separates alternate possibilities.	 

\w	Matches an alphanumeric character, including "_";  same as [A-Za-z0-9_] in ASCII, and
[\p{Alphabetic}\p{GC=Mark}\p{GC=Decimal_Number}\p{GC=Connector_Punctuation}]  

\W	Matches a non-alphanumeric character, excluding "_";
same as [^A-Za-z0-9_] in ASCII, and
[^\p{Alphabetic}\p{GC=Mark}\p{GC=Decimal_Number}\p{GC=Connector_Punctuation}]  

\s	Matches a whitespace character,
which in ASCII are tab, line feed, form feed, carriage return, and space;  

\S	Matches anything BUT a whitespace.	 

\d	Matches a digit;
same as [0-9] in ASCII;  

\D	Matches a non-digit;

^	Matches the beginning of a line or string.	  

$	Matches the end of a line or string.	  
```

Some simple regex examples 

```
 {^[-+]?[0-9]*\.?[0-9]+([eE][-+]?[0-9]+)?$}  # Floating Point Number

{^[A-Za-z]+$}   # Only letters.

 {^[[:alpha?:]]+$} # Only letters, the Unicode way.

 {(.)\1{3}} $string {\1} result # Back References 

(\[0-9]{1,3})\.(\[0-9]{1,3})\.(\[0-9]{1,3})\.(\[0-9]{1,3}) # IP Numbers 
```

# Regular Expressions in R

```{r}
letters
grep("[a-z]", letters)
grep("[x-z]", letters)
grep("[bear]", letters)
grep("[A-Z]", letters)
grep("[AB]", letters)
grep("[AB]", letters,ignore.case = TRUE)
grep("[a-zA-Z]", letters)
grep("[azAZ]", letters)
seuss <- c("You have brains in your head.",
           "You have feet in your shoes.", 
           "You can steer yourself any direction you choose.",            
           "You're on your own.", 
           "And you know what you know.", 
           "And YOU are the one who'll decide where to go...",            
           "- Dr. Seuss")
grep("you", seuss)
if(length(i <- grep("in", seuss)))
  cat("'in' appears at least once in\n\t", seuss[i], "\n")
i # 1 and 2
seuss[i]

 ## Modify all 'a' or 'b's;  "\" must be escaped 
s<-gsub("([ab])", "\\1-z\\1_s", "abc vs, ABC  - Oh, A B C. It's easy as, 1 2 3. As simple as, do re mi. a b c, 1 2 3")
s
s<-gsub("([ab])", "-z_s", "abc vs, ABC  - Oh, A B C. It's easy as, 1 2 3. As simple as, do re mi. a b c, 1 2 3")
s
s<-gsub("([ab])", "\\a\\b", "abc vs, ABC  - Oh, A B C. It's easy as, 1 2 3. As simple as, do re mi. a b c, 1 2 3")
s
stop.words <- c("The", "for", "are",
          "to",  "your", 
         "to", "and", "the", 
         "is", "your",  "and", "is",
         "it", "for", "all", "its", "it's")

seuss.re<-regexpr("[ ]+",seuss)
seuss.re
seuss.re<-gregexpr("[ ]+",seuss)
seuss.re


## trim trailing white space
str <- " You have brains in your head      "
str
str<-sub("[ ]+$", "", str)  ## trailing spaces only
str
str<-sub("^[ ]+", "", str)  ## leading spaces only
str
sub("[[:space:]]+$", "", str) ## white space, POSIX-style
sub("\\s+$", "", str, perl = TRUE) ## Perl-style white space

## capitalizing
txt <- "You have brains in your head"
gsub("(\\w)(\\w*)", "\\U\\1\\L\\2", txt, perl=TRUE)
gsub("\\b(\\w)",    "\\U\\1",       txt, perl=TRUE)

txt2 <- "You have feet in your shoes."
gsub("(\\w)(\\w*)(\\w)", "\\U\\1\\E\\2\\U\\3", txt2, perl=TRUE)
sub("(\\w)(\\w*)(\\w)", "\\U\\1\\E\\2\\U\\3", txt2, perl=TRUE)

## Decompose a URL into its components.
url <- "http://nikbearbrown/machine/learning/"
m <- regexec("^(([^:]+)://)?([^:/]+)(:([0-9]+))?(/.*)", url)
m
regmatches(url, m)
## parts of a URL:
URL_parts <- function(u) {
  m <- regexec("^(([^:]+)://)?([^:/]+)(:([0-9]+))?(/.*)", u)
  parts <- do.call(rbind,
                   lapply(regmatches(u, m), `[`, c(3L, 4L, 6L, 7L)))
  colnames(parts) <- c("protocol","host","port","path")
  parts
}
URL_parts(url)

# Spilt text
seuss
seuss[1]
sp<-strsplit(seuss[1], "a") # "You have brains in your head."
s<-"You    have   brains  in  your head. "
sp<-strsplit(seuss[1], " ") 
sp
sp<-strsplit(s," ") 
sp
sp<-strsplit(s,"[ ]+") 
sp
sp<-strsplit(s,"[ ]+",perl = TRUE) 
sp
```

Cleaning text example.  

Text Processing in R

```{r}
clean.string <- function(string){
    # Lowercase
    temp <- tolower(string)
    #' Remove everything that is not a number or letter (may want to keep more 
    #' stuff in your actual analyses). 
    temp <- stringr::str_replace_all(temp,"[^a-zA-Z\\s]", " ")
    # Shrink down to just one white space
    temp <- stringr::str_replace_all(temp,"[\\s]+", " ")
    # Split it
    temp <- stringr::str_split(temp, " ")[[1]]
    # Get rid of trailing "" if necessary
    indexes <- which(temp == "")
    if(length(indexes) > 0){
      temp <- temp[-indexes]
    } 
    return(temp)
}
clean.string(my_string)
```


```







```

